---
title: "Analyses for IRC Paper"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
This relies on the combined dfs from 1_data_import, which generates the object /sv/

```{r setup , include = FALSE}
#install.packages("kableExtra")
library(kableExtra)
library(ggplot2)
#install.packages("janitor")
library(janitor)
#install.packages("skimr")
library(skimr)
library(tidyr)
library(dplyr)
library(purrr)
```

```{r setup, include=FALSE}
# Source the script that creates `sv`
source("1_data_import.R")  # Replace with the actual script filename

# Now `sv` should exist before calling skim()
if (exists("sv")) {
  skim(sv)
}
```

```{r}

setwd("C:/Users/jessi/OneDrive - Cornell University/Documents/MSPhD/AllingtonLab_LCLUC/RStudio_WorkingDirectory")
names(sv) <- gsub("\\.x$", "", names(sv))

str(sv)
sv <- as.data.frame(sv)
sv <- sv %>% 
  mutate(across(Ref, as.factor))
view(sv)
```


Compare the numbers in hhSize to the total # reported in Camp in each age bucket
```{r}
hh_check <- sv %>% 
  select(hhSizeLarger, kidsCamp, yaCamp, adCamp, oldCamp) %>%
  rowwise() %>%
  mutate(campTot = sum(c_across(2:5), na.rm= TRUE)) %>%
# adjust campTot so that there are no zeros, min is 1
    mutate(campTot = if_else(campTot == 0, 1, campTot)) %>%
  # check the math by subtracting campTot from hhSize
    mutate(hhDif = sum(hhSizeLarger - campTot))

view(hh_check)


```

Jessie's Mods: After modifying the way we could HH members, the differences in hh_check are no longer including negative values, and the inconsistencies seem to have been nullified somewhat. 

Prior Answer: The Difference in Household size between the answer to hhSize and the total from the age buckets reported at camp ranges from -5 to 6. So clearly some respondents are indicating that hh members are they are not and vv. Basically, there are inconsistencies in their responses.  We could check further by summing across all of their answers to see how many they report in total across all locations and compare that to hh number, but ultimately we have some misunderstanding in terms of how people are answering this question. Which suggests that we might not be able to rely on it very strongly for understanding total number of people "in a hh, at camp". But that might actually be because the respondents do not have a clear delineation/conceptualization of here/gone the way we intended for the question.  Rather, some may have some hh members residing in the soum or aimag center, for school or work, but they return very frequently to the camp, so are still considered "in the household" for that answer, even if they are not reported to be living "at the camp" at the time the Q was asked.  
Also need to take into account that this survey was conducted during the winter and that may be limiting the ability of some hh members to return frequently from soum centers, UB, etc. 


#Rural out-migration dynamics
##Who is migrating?
##How has rural out-migration impacted household demographics? 

```{r}
#Extent of migration among hh: (How many hh said Yes to out-migration?)
tabyl(sv$outmigYN)
#sv %>% group_by(outmigYN) %>% count(n)

sv %>% tabyl(outmigYN, newAimag)%>%
  adorn_percentages(denominator = "col") %>%
  adorn_rounding(digits = 2)



sv %>% tabyl(concated_loc, outmigYN, ) %>%
  adorn_rounding(digits = 2) %>% 
  kable()
```
62% of the households have had a member leave the rural area.


Q: What is makeup of hh that have experienced out-migration, compared to those that have not?

A: Households who have had members migrate away have higher average # of hh members still in the camp.
```{r}
#sv %>% tabyl(outmigYN, hhSize) 
hh_summary <- sv %>% group_by(outmigYN) %>% 
  summarise(mean = mean(hhSizeLarger),
            min = min(hhSizeLarger),
            max = max(hhSizeLarger),
            sd  = sd(hhSizeLarger)) %>% 
  adorn_rounding() %>%
  kable()
view(hh_summary)

out_hh_test <- sv %>%
  group_map(~ {
    test_result <- t.test(hhSizeLarger ~ outmigYN, data = .x)
    
    # Convert to a data frame with key values
    data.frame(
      group = unique(.x$outmigYN),  # Identify the grouping variable
      statistic = test_result$statistic,
      p_value = test_result$p.value,
      estimate1 = test_result$estimate[1],  # Mean of group 1
      estimate2 = test_result$estimate[2]   # Mean of group 2
    )
  }) %>%
  bind_rows()  # Combine list into a single data frame
View(out_hh_test)


# this just yields the p.val 
sv %>%  summarise(p.value = t.test(hhSize ~ outmigYN)$p.value)
```

Let's double check that the # reported as hhSizeLarger is just those in the camp.
```{r}
sv
```

Summarize # of hh members currently living in aimag and soum centers by age bucket
```{r}
x<- sv %>% group_by(concated_loc) %>% 
  summarise(avgkidsCamp = mean(kidsCamp, na.rm = TRUE),
            avgHHyoung = mean(yaCamp, na.rm = TRUE),
            avgHHadult = mean(adCamp, na.rm = TRUE),
            avgHH60plus = mean(oldCamp, na.rm = TRUE)
            )%>% 
  adorn_rounding(digits = 2) %>%
  kable(col.names = c("", "kids", "young<br>adults", "middle<br>aged", "60+"))

```

Why hh members left:
(TEASE OUT AND BUCKET THESE ITEMS IN MORE ACCURATE CATEGORIES)
```{r}
sv %>% tabyl(mig_reason)
```



Q: What about the average age of households w/ or w/o migrants?
We can't compare actual averages bc we don't have the true age of every member, just #s in buckets.
So, we'll have to compare the distributions of hh members in the diff age buckets by mig status



A:
On average households who had indiv that had out-migrated from the rural area had more
kids, young people and adults (30-60), but fewer older people. There was no difference in the # of households in their khot Ail.
```{r}
# group by outmig and compare # hh in each, or compare distribution within each?

# compare average # of kids in the camp
sv %>% group_map(~ t.test(kidsCamp ~ outmigYN, .x))

# compare average # of 16-30 yo in the camp
sv %>% group_map(~ t.test(yaCamp ~ outmigYN, .x))

# compare average # of 30-60 yo in the camp
sv %>% group_map(~ t.test(adCamp ~ outmigYN, .x))

# compare average # of 60+ yo in the camp
sv %>% group_map(~ t.test(oldCamp ~ outmigYN, .x))

# compare average # of hh in their khot ail
sv %>% group_map(~ t.test(khotAil_num_hh ~ outmigYN, .x))

```
Contrary to our prediction, households that had had members migrate to urban areas tended to have *more* young people and *fewer* older people, compared to hh without migrants.


```{r Average number of people in each age group}

sv %>% tabyl(outmigYN, kidsCamp) 
sv %>% group_by(outmigYN) %>% 
 summarise(avgkidsCamp = mean(kidsCamp, na.rm = TRUE),
            avgHHyoung = mean(yaCamp, na.rm = TRUE),
            avgHHadult = mean(adCamp, na.rm = TRUE),
            avgHH60plus = mean(oldCamp, na.rm = TRUE)
            )%>% 
  adorn_rounding(digits = 2) %>%
  kable(col.names = c("", "kids", "young<br>adults", "middle<br>aged", "60+"))



```


Q: how do these numbers of hh size vary by district?
```{r}
sv
x<- sv %>% group_by(outmigYN, concated_loc) %>% 
  summarise(avghh = mean(hhSizeLarger, na.rm = TRUE)) 

#round values in all numeric columns to 2 decimal places
x <- x %>% mutate(across(where(is.numeric), round, 2))


x2 <- x[order(x$avghh, decreasing = TRUE),] 
kable(x2,format = "html")



```

comparing totals moved out 

```{r , echo = FALSE}
ad <- sv %>%
  select(           # select out the columns of interest here so it's not so unwieldy
   outmigYN,
   concated_loc,
   hhSizeLarger,
   kidsCamp,
   kidsSoum,
   kidsAimag,
   kidsUB,
   kidsCamp_percTotal,
   yaCamp ,
   yaSoum , 
   yaAimag , 
   yaUB ,
   adCamp,
   adSoum , 
   adAimag, 
   adUB,
   oldCamp ,
   oldSoum, 
   oldAimag , 
   oldUB, 
   hhMems_inCamp,
   hhMems_outCamp
  ) #%>% 
#  mutate(          # create new columns with totals
#    kidsOut = rowSums(across(c(kidsSoum, 
#                               kidsAimag, 
#                               kidsUB), 
#                             ~ .)),
#    yaOut = rowSums(across(c(yaSoum, 
#                             yaAimag, 
#                             yaUB), 
#                           ~ .)),
#    adOut = rowSums(across(c(adSoum,
#                             adAimag,
#                             adUB), 
#                           ~ .)),
#    oldOut = rowSums(across(c(oldSoum,
#                              oldAimag,
#                              ), 
#                            ~ .))
#  ) %>%
#  mutate(
#   totCamp = rowSums(across(c(kidsCamp,
#                               yaCamp,
#                               adCamp,
#                               oldCamp),
#                             ~.)),
#    totOut =rowSums(across(c(kidsOut,
#                             yaOut,
#                             adOut,
#                             oldOut),
#                           ~.)),
#                    )
```

Totals by location:
```{r}
ad %>%
  mutate(
    totCampPerc = if_else(hhSizeLarger > 0, round((hhMems_inCamp / hhSizeLarger) * 100, 2), 0, missing = 0),
    totOutPerc = if_else(hhSizeLarger > 0, round((hhMems_outCamp / hhSizeLarger) * 100, 2), 0, missing = 0)
  ) %>%
  group_by(outmigYN) %>%
  summarise(
    across(c(hhSizeLarger, hhMems_inCamp, hhMems_outCamp, totCampPerc, totOutPerc), ~ round(mean(.x, na.rm = TRUE), 2))
  )

ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(hhSizeLarger, hhMems_inCamp, hhMems_outCamp), sum, na.rm = TRUE)
  )

```
Jessie's Mods: Not only are the households who say that members of the HH live in the soum/aimag/UB larger, a larger percentage of the household has outmigrated compared to HHs who say that HH members don't live in the soum/aimag/UB. I'm not sure what this would mean, maybe that everyone has decided that regardless of HH size, the number of people needed to do herding labor is the same, which means that larger HH's send a greater percentage of people away? I guess we would need to see if everyone is complaining about the same issues regardless of HH members in camp.  

```{r}
ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(kidsCamp, kidsSoum, kidsAimag, kidsUB), ~ round(mean(.x, na.rm = TRUE), 2))
  ) 

ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(kidsCamp, kidsSoum, kidsAimag, kidsUB), sum, na.rm = TRUE)
  ) 

```
```{r}
ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(yaCamp, yaSoum, yaAimag, yaUB), ~ round(mean(.x, na.rm = TRUE), 2))
  ) 

ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(yaCamp, yaSoum, yaAimag, yaUB), sum, na.rm = TRUE)
  ) 
```
```{r}
ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(adCamp, adSoum, adAimag, adUB), ~ round(mean(.x, na.rm = TRUE), 2))
  ) 

ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(adCamp, adSoum, adAimag, adUB), sum, na.rm = TRUE)
  ) 
```

```{r}
ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(oldCamp, oldSoum, oldAimag, oldUB), ~ round(mean(.x, na.rm = TRUE), 2))
  ) 

ad %>% 
  group_by(outmigYN) %>%
  summarise(
    across(c(oldCamp, oldSoum, oldAimag, oldUB), sum, na.rm = TRUE)
  ) 
```

#Why moved
```{r}

```


